### 2013-03-26 doc/sosp13/abstract.tex
changes to abstract & first 3 paragraphs of intro

\Grappa is a runtime system for commodity clusters of multicore computers that
presents a massively parallel, single address space abstraction to
applications. \Grappa's goal is to provide scalable performance for irregular
parallel applications, such as graph processing. These applications are
challenging because they exhibit little data locality, imbalanced parallel
work and complex communication patterns.

\Grappa serves both as a C++ user library and as a foundation upon which
higher level languages can be developed or adapted. \Grappa tolerates delays
to remote memory by multiplexing thousands of lightweight workers
to each processor core; balances load via fine-grained distributed
work-stealing; takes full advantage of network characteristics by aggregating
smaller data requests into large ones; and provides efficient synchronization
and remote operations. We present a detailed description of the \Grappa
system,
%programming examples using the library interface,
and compare
performance for several irregular benchmarks to hand-optimized MPI code and
the Cray XMT, a custom system used to target the real time graph analytics
market.

>>>>

\Grappa is a runtime system for commodity clusters of multicore computers that
presents a massively parallel, single address space abstraction to
applications. \Grappa's purpose is scalable performance for irregular
parallel applications, such as graph processing. Poor data locality, imbalanced parallel
work and complex communication patterns make these applications difficult to scale.

\Grappa serves both as a C++ user library and as a foundation upon which
higher level languages can be developed or adapted. \Grappa tolerates delays
to remote memory by multiplexing thousands of lightweight workers
to each processor core; balances load via fine-grained distributed
work-stealing; increases communication throughput by aggregating
smaller data requests into large ones; and provides efficient synchronization
and remote operations. We present a detailed description of the \Grappa
system and
%programming examples using the library interface,
performance comparisons on several irregular benchmarks to hand-optimized MPI code and to
the Cray XMT, a custom system used to target the real time graph analytics
market.

### 2013-03-26 doc/sosp13/intro.tex
execution contexts to workers
changes to abstract & first 3 paragraphs of intro

Irregular applications exhibit workloads, dependences, and memory accesses
that are highly sensitive to input. Classic examples of such applications
include branch and bound optimization, SPICE circuit simulation, contact
algorithms in car crash analysis, and network flow. Important contemporary
examples include processing large graphs in the business, national security,
machine learning, and social network computing domains. For these emerging
applications, fast response -- given the sheer amount of data -- requires
multinode systems. The most broadly available multinode systems are those
built from x86 compute nodes interconnected via ethernet or InfiniBand.
Scalable performance of irregular applications on these systems is challenging
for two reasons:

\noindent{\bf Irregular applications exhibit little data locality.} Data
accesses tend to be spread across the entire memory of the system. This makes
current memory hierarchy features ineffective. Caches are of little assistance
with such low data re-use and spatial locality. Commodity prefetching hardware
is effective only when addresses are known many cycles before the data is
consumed or the accesses follow a predictable pattern, neither of which occurs
in irregular applications. As a consequence, commodity microprocessors stall
often when executing irregular applications and data-parallel distributed
frameworks (e.g., MapReduce~\cite{mapreduce:osdi04}) are ineffective.

\noindent{\bf Irregular applications frequently request small amounts of
off-node data.} On multi-node systems, the challenges presented by low
locality are analogous, and exacerbated by the increased latency of going
off-node. Irregular applications also present a challenge to mass market
network technology, which is designed to transfer large blocks of data, not
the word-sized references emitted by irregular application tasks. Injection
rate into the network is insufficient to utilize wire bandwidth when blocks
are below about two-kilobytes, so any naive communication strategy severely
under-utilizes the network.

@@

We believe we can support irregular applications with good performance and
cost-effectiveness with commodity hardware for two main reasons. First,
commodity multicore processors have become extremely fast and cheap. So with
careful software engineering we can multiplex hundred of thousands of
execution contexts into a single core. Second, commodity networks offer high
bandwidth as long as messages are large enough. So with enough data requests
in flight simultaneously, the system can aggregate small messages into large
enough ones.

>>>>

Irregular applications exhibit workloads, dependences, and memory accesses
that are highly sensitive to input. Classic examples of such applications
include branch and bound optimization, SPICE circuit simulation, contact
algorithms in car crash analysis, and network flow. Important contemporary
examples include processing large graphs in the business, national security,
machine learning, and social network computing domains. For these emerging
applications, fast response -- given the sheer amount of data -- requires
multinode systems. The most broadly available multinode systems are those
built from x86 compute nodes interconnected via ethernet or InfiniBand.
Scalable performance of irregular applications on these systems is elusive
for two reasons:

\noindent{\bf Low Injection Rate} Reference patterns of irregular
applications are unpredictable, with odds increasing as data is
spread out more broadly across a larger system that a given reference
will refer to memory located remotely.  This behavior constrasts
sharply with that of data-parallel applications for which efficient
frameworks such as MapReduce~\cite{mapreduce:osdi04} may exploit the
structure of computation and data affinity to limit internode
communication to comparably infrequent and larger messages.  Mass
market network technology is designed to transfer these larger blocks
of data, not the word-sized references emitted by irregular
application tasks. Injection rate into the network is insufficient to
utilize wire bandwidth when blocks are below about two-kilobytes, so
any naive communication strategy severely under-utilizes the network.

\noindent{\bf High Latency} Unlike data parallel applications, whose
communications are regular and expected, and can therefore easily be
overlapped with computation at the application level, irregular
application tasks encounter remote references throughout their execution
and must resolve them before making further progress.  The latency of
commodity networks is anywhere from a one to hundreds of microseconds,
far more than can be tolerated via the simultaneous multithreading
available in x86 processors.  Consequently, even with RDMA, processor
prefetching and caching hardware alone is inadequate to ensure pipeline
utilization in the presence of high latency networks.

@@

We believe we can support irregular applications with good performance and
cost-effectiveness with commodity hardware for two main reasons. First,
commodity multicore processors have become extremely fast and cheap. So with
careful software engineering we can multiplex hundred of thousands of
workers into a single core. Second, commodity networks offer high
bandwidth as long as messages are large enough. So with enough data requests
in flight simultaneously, the system can aggregate small messages into large
enough ones.
